---
title: "Week 14 IP Part 3 - Supermarket_Sales_Forecasting - Sales"
output:
  pdf_document: default
  html_notebook: default
---

```{r}
install.packages("pacman")
# Loading the necessary libraries and automatically installing them if not present
pacman :: p_load(rio,tidyverse, Amelia ,anomalize)
```
```{r}
#Load the readr library to bring in the dataset
library(readr)
```

```{r}


#Reading and checking the sales data
data <- read.csv('http://bit.ly/CarreFourSalesDataset')
head(data)
tail(data)

```



```{r}
# Checking the size and shape of data
dim(data)
```
We have 1000 rows and 2 columns
```{r}
# Viewing data types using str().
str(data)
```

```{r}
#Checking on summary statistics of numeric variables

summary(data)
```
##3. Data Cleaning 
```{r}
## To ensure uniformity, I will lowercase all the columns
names(data)<- tolower(names(data))
head(data) 
```
```{r}

#changing date to date time. 
data$date <- as.Date(data$date, "%m/%d/%y")
head(data)
```



Next I'm going to checking for missing values in our dataset,,Missing values may affect the perfomance of our model, so we will find a way to deal with them

```{r}
#Checking for missing values in the columns of our dataset
#

colSums(is.na(data))

```
There is no missing data in any column
Lets now check for duplicates in our dataset
```{r}
anyDuplicated(data)
```
We dont have duplicates in our dataset
Lets now check for outliers in our dataset
```{r}
ggplot(data) +
  aes(x = "", y = sales) +
  geom_boxplot(fill = "#0c4c8a") +
  theme_minimal()

```
There are afew outliers where sales exceed 950
We will not drop them.
### Detecting our anomalies



### checking  the number of transaction in a day

```{r}
# group and tally the number of transactions per day
df <- data %>% group_by(date) %>% tally()
colnames(df) <- c('transactionDate', 'totalCount')
head(df)
```
the above shows the number of the customer transactions frm Jan 1st to Jan 6th. 
```{r}
# we now plot using plot_anomaly_decomposition() to visualize out data.
df %>%
    time_decompose(totalCount) %>%
    anomalize(remainder) %>%
    plot_anomaly_decomposition(ncol = 1, alpha_dots = 0.5) +
    ggtitle("Anomaly Detection Plot")
```
observation: there is no red flag thus these customer transactions do not have anomalies.

```{r}
# ploting the recomposition to try and see anomalies
#
df %>%
    time_decompose(totalCount) %>%
    anomalize(remainder) %>%
    time_recompose() %>%
    plot_anomalies(time_recomposed = TRUE, ncol = 1, alpha_dots = 0.25, fill="pink") +
    ggtitle("Anomalie detection plots")
```

observation: we can conclude that there is no anomaly(outliers) in this dataset.

